---
title: "Logging Beyond Local Files"
date: "Updated as of `r as.Date(Sys.time())`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Logging Beyond Local Files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(log4r)
```

For local development or simple batch R scripts run manually, writing log
messages to a file for later inspection (with `file_appender`) is quite
convenient. However, for deployed R applications (like Shiny apps and Plumber
APIs) or automated scripts it is more likely that all an organization's logs
will be aggregated in one central place (perhaps with a commercial tool or
service[^1]) for searching and monitoring. It can be annoying or impossible to
upload log files in these cases.

If your organization's platform supports reading log messages from regular
program output,[^2] you can just use the default setup, which uses the
`console_appender()`. Otherwise, **log4r** includes three additional appenders
to facilitate shipping logs off to an aggregator:

* `syslog_appender`: For writing messages to the system log on Linux, macOS, and
  other Unix-like operating systems.

* `http_appender`: For sending log messages as HTTP requests.

* `tcp_appender`: For writing log messages to TCP connections.

## Writing to the System Log

The Unix "System log" (syslog) dates to the mid-1980s, and is still widely used.
Almost all log aggregation services support ingesting a server's syslog
messages, so often the easiest way to get your logs to these services is to make
your R talk to the local syslog.

To use the `syslog_appender`, all you need is an identifier for your R app or
script:

```{r rsyslog, eval=requireNamespace("rsyslog", quietly = TRUE)}
logger <- logger(appenders = syslog_appender("my-R-script"))
```

Requires the [**rsyslog**](https://cran.r-project.org/package=rsyslog) package.

## Sending Logs over HTTP

If you're not already forwarding syslog messages (or need to send logs from
Windows), the next most-common approach is to send them over HTTP. Log
aggregation services usually provide an HTTP API endpoint to facilitate this:

```{r http-1}
logger <- logger(appenders = http_appender("http://logging.example.local"))
```

Some services use `GET` or `PUT` requests instead of the more intuitive `POST`,
which you can opt into as follows:

```{r http-2}
logger <- logger(
  appenders = http_appender("http://logging.example.local", method = "GET")
)
```

Finally, if you need complete control over the HTTP request (for example, to
send a specific header or use authentication), you can pass additional
parameters to the underlying **httr** verb function:

```{r http-3}
logger <- logger(
  appenders = http_appender(
    "http://logging.example.local",
    method = "GET",
    layout = default_log_layout(),
    httr::add_headers(`X-Custom-Header` = 1),
    httr::user_agent("my-r-script/1.0.0")
  )
)
```


Requires the [**httr**](https://cran.r-project.org/package=httr) package.

## Writing Directly to TCP Connections

For some workloads, the send-and-receive structure of HTTP requests may be
undesirable, so many log aggregators also accept messages directly at a TCP
port:

```{r tcp, eval = FALSE}
logger <- logger(
  appenders = tcp_appender("tcp://logging.example.local", port = 9551)
)
```

[^1]: Such as [Splunk](https://www.splunk.com/) or
[Loggly](https://www.loggly.com/solution/log-management/). There is also are
also many open-source options, such as the [ELK
Stack](https://www.elastic.co/what-is/elk-stack) or
[Graylog](https://www.graylog.org/).

[^2]: For example, when running as a Docker container in a cluster with all logs
forwarded automatically, or when scripts are wrapped up as SystemD unit files.
